---
title: "Coffee Data"
author: "Katelyn Donaty"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Visualization
## Exploring and visualizing data by answering the following questions:
* How have sales trended over time?
* Which days of the week tend to be busiest?
* Which products are sold the most/least often?
* Which products drive the most revenue? 
* What is the most popular brewed coffee?
* What is the most popular brewed tea?
* What is the most popular bakery item?

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(modelr)
# library(readxl)
library(dplyr)
library(ggplot2)
library(lubridate)
library(randomForest)
library(e1071)
```



```{r, echo=FALSE}
#coffee_data <- read_excel("Coffee Shop Sales.xlsx")
#write_csv(coffee_data, "Coffee Shop Sales.csv")
coffee_data_csv <- read.csv("Coffee Shop Sales.csv")
df <- data.frame(coffee_data_csv)
# df
```


## Checking for NA values
```{r, echo=FALSE}
na <- sum(is.na(df))
na
```
No NA values were found.     

# Summary stats
```{r, echo=FALSE}
summary(df)
```

## Store locations
```{r, echo=FALSE}
unique(df$store_location)
```

## Product categories
```{r, echo=FALSE}
unique(df$product_category)

```


## Coffee sales over time (Lower Manhattan store)
```{r, fig.width=7, fig.height=4, echo=FALSE}
lower_manhattan <- subset(df, select = c(store_location, transaction_date, product_category, unit_price))

lower_manhattan <- lower_manhattan %>%
  mutate(transaction_date = as.Date(transaction_date))

lower_manhattan_focused <- lower_manhattan %>%
  filter(startsWith(store_location, "Lower Manhattan")) %>%
  filter(startsWith(product_category, "Coffee")) %>%
  group_by(transaction_date) %>%
  summarise(total_sales = sum(unit_price, na.rm=TRUE))

dates <- sort(lower_manhattan_focused$transaction_date)
first <- dates[1]
middle <- dates[floor(length(dates) / 2)]
last <- dates[length(dates)]
labels <- c(first, middle, last)

lower_manhattan_coffee <- ggplot(lower_manhattan_focused, aes(x = transaction_date, y = total_sales, group=1)) +
  geom_line() +
  labs(title = "Coffee Sales Over Time (Lower Manhattan)", x = "Date", y = "Sales")

#lower_manhattan_focused
lower_manhattan_coffee
```

## Bakery sales over time (Lower Manhattan store)
```{r, fig.width=7, fig.height=4, echo=FALSE}
lower_manhattan <- subset(df, select = c(store_location, transaction_date, product_category, unit_price))

lower_manhattan <- lower_manhattan %>%
  mutate(transaction_date = as.Date(transaction_date))

lower_manhattan_focused <- lower_manhattan %>%
  filter(startsWith(store_location, "Lower Manhattan")) %>%
  filter(startsWith(product_category, "Bakery")) %>%
  group_by(transaction_date) %>%
  summarise(total_sales = sum(unit_price, na.rm=TRUE))

lower_manhattan_bakery <- ggplot(lower_manhattan_focused, aes(x = transaction_date, y = total_sales)) +
  geom_line() +
  labs(title = "Bakery Sales Over Time (Lower Manhattan)", x = "Date", y = "Sales")

#lower_manhattan_focused
lower_manhattan_bakery
```

## Total sales over time (Lower Manhattan store)
```{r, fig.width=7, fig.height=4, echo=FALSE}
lower_manhattan <- subset(df, select = c(store_location, transaction_date, unit_price))

lower_manhattan <- lower_manhattan %>%
  mutate(transaction_date = as.Date(transaction_date))

lower_manhattan_focused <- lower_manhattan %>%
  filter(startsWith(store_location, "Lower Manhattan")) %>%
  group_by(transaction_date) %>%
  summarise(total_sales = sum(unit_price, na.rm=TRUE))

lower_manhattan_total <- ggplot(lower_manhattan_focused, aes(x = transaction_date, y = total_sales)) +
  geom_line() +
  labs(title = "Total Sales Over Time (Lower Manhattan)", x = "Date", y = "Sales")

#lower_manhattan_focused
lower_manhattan_total
```

## Total sales over time (all stores)
```{r, fig.width=7, fig.height=4, echo=FALSE}
all_stores <- subset(df, select = c(transaction_date, unit_price))

all_stores <- all_stores %>%
  mutate(transaction_date = as.Date(transaction_date))

all_stores_focused <- all_stores %>%
  group_by(transaction_date) %>%
  summarise(total_sales = sum(unit_price, na.rm=TRUE))

all_stores_total <- ggplot(all_stores_focused, aes(x = transaction_date, y = total_sales)) +
  geom_line() +
  labs(title = "Total Sales Over Time (All Stores)", x = "Date", y = "Sales")

#all_stores_focused
all_stores_total
```

## Total sales per month (all stores)
```{r, fig.width=7, fig.height=4, echo=FALSE}
all_stores <- subset(df, select = c(transaction_date, unit_price))

all_stores <- all_stores %>%
  mutate(transaction_date = as.Date(transaction_date))

all_stores_monthly <- all_stores %>%
  mutate(month = floor_date(transaction_date, unit = "month")) %>%
  group_by(month) %>%
  summarise(total_sales = sum(unit_price, na.rm = TRUE))

all_stores_monthly_total <- ggplot(all_stores_monthly, aes(x = month, y = total_sales, fill = factor(month))) +
  geom_col() +
  labs(title = "Total Sales Per Month (All Stores)", x = "Month", y = "Sales", fill = "Total Sales Per Month (All Stores)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#CB997E", "#DDBEA9", "#FFE8D6", "#B7B7A4", "#A5A58D", "#6B705C"),
    labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun")
    )

#all_stores_monthly
all_stores_monthly_total
```

## Total sales per day of the week (all stores)     
Mondays and Fridays are the busiest overall     
```{r, fig.width=7, fig.height=4, echo=FALSE}
all_stores <- subset(df, select = c(transaction_date, unit_price))

all_stores <- all_stores %>%
  mutate(transaction_date = as.Date(transaction_date))

all_stores_weekday <- all_stores %>%
  mutate(weekday = wday(transaction_date, label = TRUE, abbr = FALSE)) %>%
  group_by(weekday) %>%
  summarise(total_sales = sum(unit_price, na.rm = TRUE))

all_stores_weekday_total <- ggplot(all_stores_weekday, aes(x = weekday, y = total_sales, fill = factor(weekday))) +
  geom_col() +
  labs(title = "Total Sales Per Day of the Week (All Stores)", x = "Weekday", y = "Sales", fill = "Total Sales Per Day of the Week (All Stores)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#797D62", "#9B9B7A", "#D9AE94", "#F1DCA7", "#FFCB69", "#D08C60", "#997B66"),
    labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")
    ) +
  scale_x_discrete(labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))

#all_stores_weekday
all_stores_weekday_total
```

## Total sales per day of the week (Lower Manhattan)     
Mondays and Fridays are the busiest     
```{r, fig.width=7, fig.height=4, echo=FALSE}
lower_manhattan <- subset(df, select = c(store_location, transaction_date, unit_price))

lower_manhattan <- lower_manhattan %>%
  mutate(transaction_date = as.Date(transaction_date))

lower_manhattan_weekday <- lower_manhattan %>%
  mutate(weekday = wday(transaction_date, label = TRUE, abbr = FALSE)) %>%
  filter(startsWith(store_location, "Lower Manhattan")) %>%
  group_by(weekday) %>%
  summarise(total_sales = sum(unit_price, na.rm = TRUE))

lower_manhattan_weekday_total <- ggplot(lower_manhattan_weekday, aes(x = weekday, y = total_sales, fill = factor(weekday))) +
  geom_col() +
  labs(title = "Total Sales Per Day of the Week (Lower Manhattan)", x = "Weekday", y = "Sales", fill = "Total Sales Per Day of the Week (Lower Manhattan)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#797D62", "#9B9B7A", "#D9AE94", "#F1DCA7", "#FFCB69", "#D08C60", "#997B66"),
    labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")
    ) +
  scale_x_discrete(labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))

#lower_manhattan_weekday
lower_manhattan_weekday_total
```

## Best product categories for revenue (all stores)     
Coffee and Tea (not Loose Tea) bring in the most profit, followed by Bakery items     
```{r, fig.width=7, fig.height=4, echo=FALSE}
all_stores <- subset(df, select = c(transaction_date, product_category, unit_price))
all_stores_focused <- all_stores %>%
  group_by(product_category) %>%
  summarise(total_sales = sum(unit_price, na.rm = TRUE))

all_stores_total <- ggplot(all_stores_focused, aes(x = product_category, y = total_sales, fill = factor(product_category))) +
  geom_col() +
  labs(title = "Best Products for Revenue (All Stores)", x = "Product Category", y = "Sales", fill = "Best Products for Revenue (All Stores)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#B98B73", "#CB997E", "#DDBEA9", "#FFE8D6", "#D4C7B0", "#B7B7A4", "#A5A58D", "#6B705C", "#3F4238"),
    labels = c("Bakery", "Branded", "Coffee", "Coffee Beans", "Drinking Chocolate", "Flavors", "Loose Tea", "Packaged Chocolate", "Tea")
    ) +
  scale_x_discrete(labels = c("Bakery", "Branded", "Coffee", "Coffee Beans", "Drinking Chocolate", "Flavors", "Loose Tea", "Packaged Chocolate", "Tea")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#all_stores_focused
all_stores_total
```

## Best-selling products (all stores)     
Coffee and Tea (not Loose Tea) sells the most, followed by Bakery items     
```{r, fig.width=7, fig.height=4, echo=FALSE}
all_stores <- subset(df, select = c(transaction_date, product_category, transaction_qty))
all_stores_focused <- all_stores %>%
  group_by(product_category) %>%
  summarise(total_amt = sum(transaction_qty, na.rm = TRUE))

all_stores_total <- ggplot(all_stores_focused, aes(x = product_category, y = total_amt, fill = factor(product_category))) +
  geom_col() +
  labs(title = "Best-Selling Product Categories (All Stores)", x = "Product Category", y = "Amount", fill = "Best-Selling Product Categories (All Stores)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#B98B73", "#CB997E", "#DDBEA9", "#FFE8D6", "#D4C7B0", "#B7B7A4", "#A5A58D", "#6B705C", "#3F4238"),
    labels = c("Bakery", "Branded", "Coffee", "Coffee Beans", "Drinking Chocolate", "Flavors", "Loose Tea", "Packaged Chocolate", "Tea")
    ) +
  scale_x_discrete(labels = c("Bakery", "Branded", "Coffee", "Coffee Beans", "Drinking Chocolate", "Flavors", "Loose Tea", "Packaged Chocolate", "Tea")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#all_stores_focused
all_stores_total
```


## Busiest hours (Monday, all stores)
```{r, fig.width=7, fig.height=4, echo=FALSE}
#unique(df$transaction_time)[1:50]

monday_hourly_df <- df %>%
  mutate(
    time_parsed = ymd_hms(transaction_time, tz="UTC"),
    hour = hour(time_parsed)
  )

ggplot(monday_hourly_df, aes(x = hour)) +
  geom_density(fill = "#CB997E", alpha=0.6) +
  labs(title = "Sales Per Hour on Mondays", x = "Hour of Day", y = "Sales") +
  theme_minimal()

```

## Sales per store
```{r, fig.width=7, fig.height=4, echo=FALSE}
all_stores <- df %>%
  select(store_location, unit_price) %>%
  group_by(store_location) %>%
  summarise(total_sales = sum(unit_price, na.rm = TRUE))

astoria_percent <- all_stores %>%
  filter(startsWith(store_location, "Astoria"))
astoria_percent <- astoria_percent$total_sales/sum(all_stores$total_sales)

lower_manhattan_percent <- all_stores %>%
  filter(startsWith(store_location, "Lower Manhattan"))
lower_manhattan_percent <- lower_manhattan_percent$total_sales/sum(all_stores$total_sales)

hells_kitchen_percent <- all_stores %>%
  filter(startsWith(store_location, "Hell's Kitchen"))
hells_kitchen_percent <- hells_kitchen_percent$total_sales/sum(all_stores$total_sales)

sales_per_store_df <- data.frame(
  item = c("Astoria", "Lower Manhattan", "Hell's Kitchen"),
  value = c(astoria_percent, lower_manhattan_percent, hells_kitchen_percent)
)

sales_per_store_df <- sales_per_store_df %>%
  mutate(
    percent = value/sum(value) * 100,
    label = paste0(item, " (", round(percent, 2), "%")
  )

ggplot(sales_per_store_df, aes(x = "", y = value, fill = label)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta="y") +
  labs(title = "Sales Per Store", fill = "Sales Per Store") +
  theme_void() + 
  scale_fill_manual(values = c("#CB997E", "#D4C7B0", "#6B705C"),
                    breaks = sales_per_store_df$label,
                    labels = sales_per_store_df$label)

```

## Most popular coffee across all stores
```{r, fig.width=7, fig.height=4, echo=FALSE}
coffee <- subset(df, select = c(product_category, product_type, transaction_qty))
coffee_brewed <- coffee %>%
  filter(str_detect(product_type, regex("Brewed|Barista Espresso|Drip Coffee", ignore_case = TRUE)), product_category == "Coffee") %>%
  group_by(product_type) %>%
  summarise(total_amt = sum(transaction_qty, na.rm = TRUE)) %>%
  arrange(desc(total_amt))
#coffee_brewed

coffee_total <- ggplot(coffee_brewed, aes(x = reorder(product_type, -total_amt), y = total_amt, fill = factor(product_type))) +
  geom_col() +
  labs(title = "Best-Selling Coffee (All Stores)", x = "Coffee Type", y = "Amount", fill = "Best-Selling Coffee (All Stores)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#DDBEA9", "#FFE8D6", "#D4C7B0", "#B7B7A4", "#A5A58D"),
    labels = c("Barista Espresso", "Drip Coffee", "Gourmet Brewed Espresso", "Organic Brewed Espresso", "Premium Brewed Espresso")
    ) +
  scale_x_discrete(labels = c("Gourmet Brewed Coffee", "Barista Espresso", "Organic Brewed Coffee", "Drip Coffee", "Premium Brewed Coffee")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
coffee_total
```


## Most popular tea
```{r, fig.width=7, fig.height=4, echo=FALSE}
tea <- subset(df, select = c(product_category, product_type, transaction_qty))
tea_brewed <- tea %>%
  filter(startsWith(product_category, "Tea")) %>%
  group_by(product_type) %>%
  summarise(total_amt = sum(transaction_qty, na.rm = TRUE)) %>%
  arrange(desc(total_amt))
#tea_brewed

tea_total <- ggplot(tea_brewed, aes(x = reorder(product_type, -total_amt), y = total_amt, fill = factor(product_type))) +
  geom_col() +
  labs(title = "Best-Selling Tea (All Stores)", x = "Tea Type", y = "Amount", fill = "Best-Selling Tea (All Stores)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#DDBEA9", "#D4C7B0", "#B7B7A4", "#A5A58D"),
    labels = c("Brewed Black Tea", "Brewed Chai Tea", "Brewed Green Tea", "Brewed Herbal Tea")
    ) +
  scale_x_discrete(labels = c("Brewed Chai Tea", "Brewed Black Tea", "Brewed Herbal Tea", "Brewed Green Tea")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
tea_total
```

## Most popular bakery
```{r, fig.width=7, fig.height=4, echo=FALSE}
bakery <- subset(df, select = c(product_category, product_type, transaction_qty))
bakery_items <- bakery %>%
  filter(startsWith(product_category, "Bakery")) %>%
  group_by(product_type) %>%
  summarise(total_amt = sum(transaction_qty, na.rm = TRUE)) %>%
  arrange(desc(total_amt))
#bakery_items

bakery_total <- ggplot(bakery_items, aes(x = reorder(product_type, -total_amt), y = total_amt, fill = factor(product_type))) +
  geom_col() +
  labs(title = "Best-Selling Bakery Items (All Stores)", x = "Item Type", y = "Amount", fill = "Best-Selling Bakery Items (All Stores)") +
  theme_minimal() +
  scale_fill_manual(
    values = c("#DDBEA9", "#D4C7B0", "#B7B7A4"),
    labels = c("Biscotti", "Pastry", "Scone")
    ) +
  scale_x_discrete(labels = c("Scone", "Pastry", "Biscotti")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
bakery_total
```





# Predictions Using Linear Regression, Random Forest, and Support Vector Machines
## Predicting future data by answering the following questions:
* What is the expected coffee sales trend for July to December?
* What is the expected total revenue for the end of December?

## Coffee Sales for July-December (Linear Regression)
```{r, echo=FALSE}
predicted_coffee_sales_data <- df %>%
  select(transaction_date, unit_price) %>%
  mutate(transaction_date = as.Date(transaction_date)) %>%
  mutate(date_numeric = as.numeric(transaction_date)) %>%
  group_by(transaction_date) %>%
  summarise(total_sales = sum(unit_price, ra.rm=TRUE))

model <- lm(total_sales ~ transaction_date, data = predicted_coffee_sales_data)

last_date <- max(predicted_coffee_sales_data$transaction_date)
future_dates <- seq.Date(from = last_date + 1, by = "day", length.out = 184)

future_data <- data.frame(
  transaction_date = future_dates,
  date_numeric = as.numeric(future_dates)
)

future_data$predicted_sales <- predict(model, newdata = future_data)

combined_data <- bind_rows(
  predicted_coffee_sales_data %>% select(transaction_date, total_sales) %>% mutate(type="Actual"),
           future_data %>% rename(total_sales = predicted_sales) %>% mutate(type = "Predicted")
           )
```

## Linear Regression Model Summary
```{r, echo=FALSE}
summary(model)
```


## Actual and Predicted Coffee Sales (Linear Regression)
```{r, fig.width=7, fig.height=4, echo=FALSE}
ggplot(combined_data, aes(x = transaction_date, y = total_sales, color = type)) +
  geom_line() +
  labs(title = "Actual and Predicted Coffee Sales (Linear Regression)", x = "Date", y = "Sales") +
  theme_minimal() +
  scale_color_manual(
    values = c("Actual" = "#CB997E", "Predicted" = "#A5A58D")
    )
```

## Predicted total revenue (Linear Regression)
Dec 31, 2023: $1,514,940     
Jan 1, 2023: $1,732.80
```{r, echo=FALSE}
predicted_coffee_sales_data <- df %>%
  select(transaction_date, unit_price) %>%
  mutate(transaction_date = as.Date(transaction_date)) %>%
  mutate(date_numeric = as.numeric(transaction_date)) %>%
  group_by(transaction_date) %>%
  summarise(total_sales = sum(unit_price, ra.rm=TRUE))

model <- lm(total_sales ~ transaction_date, data = predicted_coffee_sales_data)

last_date <- max(predicted_coffee_sales_data$transaction_date)
future_dates <- seq.Date(from = last_date + 1, by = "day", length.out = 184)

future_data <- data.frame(
  transaction_date = future_dates,
  date_numeric = as.numeric(future_dates)
  )

future_data$predicted_sales <- predict(model, newdata = future_data)

combined_data <- bind_rows(
  predicted_coffee_sales_data %>% select(transaction_date, total_sales) %>% mutate(type="Actual"),
           future_data %>% rename(total_sales = predicted_sales) %>% mutate(type = "Predicted")
           )

total_revenue <- combined_data %>%
  summarise(expected_total_revenue = sum(total_sales, na.rm = TRUE))
#total_revenue
```

## Coffee Sales for July-December (Random Forest)
```{r, echo=FALSE}
rf_data <- df %>%
  mutate(
    transaction_date = as.Date(transaction_date),
    date_numeric = as.numeric(transaction_date),
    day_of_week = wday(transaction_date, label = TRUE),
    month = month(transaction_date, label = TRUE),
  ) %>%
  group_by(transaction_date) %>%
  summarise(
    total_sales = sum(unit_price, na.rm = TRUE),
    date_numeric = first(date_numeric),
    day_of_week = first(day_of_week),
    month = first(month)
  ) %>%
  ungroup()

set.seed(444)

rf_model <- randomForest(total_sales ~ date_numeric + day_of_week + month, data = rf_data, ntree = 500)

rf_data$predicted_sales <- predict(rf_model, newdata = rf_data)

last_date <- max(rf_data$transaction_date)
future_dates <- seq.Date(last_date + 1, by = "day", length.out = 184)

future_data <- data.frame (
  transaction_date = future_dates,
  date_numeric = as.numeric(future_dates),
  day_of_week = wday(future_dates, label = TRUE),
  month = month(future_dates, label = TRUE)
)

future_data$predicted_sales <- predict(rf_model, newdata = future_data)

combined_data <- bind_rows(
  rf_data %>% select(transaction_date, total_sales) %>% mutate(type="Actual"),
           future_data %>% rename(total_sales = predicted_sales) %>% mutate(type = "Predicted")
           )
```

## Random Forest Model Summary
```{r, echo=FALSE}
summary(rf_model)
```


## Actual and Predicted Coffee Sales (Random Forest)
```{r, fig.width=7, fig.height=4, echo=FALSE}
ggplot(combined_data, aes(x = transaction_date, y = total_sales, color = type)) +
  geom_line() +
  labs(title = "Actual and Predicted Coffee Sales (Random Forest)", x = "Date", y = "Sales", color = "Actual and Predicted Coffee Sales") +
  theme_minimal() +
  scale_color_manual(
    values = c("Actual" = "#CB997E", "Predicted" = "#A5A58D")
    )
```

## Predicted total revenue (Random Forest)
Dec 31, 2023: $1,216,976     
Jan 1, 2023: $1,732.80
```{r, echo=FALSE}
rf_data <- df %>%
  mutate(
    transaction_date = as.Date(transaction_date),
    date_numeric = as.numeric(transaction_date),
    day_of_week = wday(transaction_date, label = TRUE),
    month = month(transaction_date, label = TRUE),
  ) %>%
  group_by(transaction_date) %>%
  summarise(
    total_sales = sum(unit_price, na.rm = TRUE),
    date_numeric = first(date_numeric),
    day_of_week = first(day_of_week),
    month = first(month)
  ) %>%
  ungroup()

set.seed(444)

rf_model <- randomForest(total_sales ~ date_numeric + day_of_week + month, data = rf_data, ntree = 500)

rf_data$predicted_sales <- predict(rf_model, newdata = rf_data)

last_date <- max(rf_data$transaction_date)
future_dates <- seq.Date(last_date + 1, by = "day", length.out = 184)

future_data <- data.frame (
  transaction_date = future_dates,
  date_numeric = as.numeric(future_dates),
  day_of_week = wday(future_dates, label = TRUE),
  month = month(future_dates, label = TRUE)
)

future_data$predicted_sales <- predict(rf_model, newdata = future_data)

combined_data <- bind_rows(
  rf_data %>% select(transaction_date, total_sales) %>% mutate(type="Actual"),
           future_data %>% rename(total_sales = predicted_sales) %>% mutate(type = "Predicted")
           )

total_revenue <- combined_data %>%
  summarise(expected_total_revenue = sum(total_sales, na.rm = TRUE))
#total_revenue
```

## Coffee Sales for July-December (SVM)
```{r, echo=FALSE}
svm_data <- df %>%
  mutate(
    transaction_date = as.Date(transaction_date),
    date_numeric = as.numeric(transaction_date),
    day_of_week = wday(transaction_date, label = TRUE),
    month = month(transaction_date, label = TRUE),
  ) %>%
  group_by(transaction_date) %>%
  summarise(
    total_sales = sum(unit_price, na.rm = TRUE),
    date_numeric = first(date_numeric),
    day_of_week = first(day_of_week),
    month = first(month)
  ) %>%
  ungroup()

svm_model <- svm(total_sales ~ date_numeric + day_of_week + month, data = svm_data, type = "eps-regression")

svm_data$predicted_sales <- predict(svm_model, newdata = svm_data)

last_date <- max(svm_data$transaction_date)
future_dates <- seq.Date(last_date + 1, by = "day", length.out = 184)

future_data <- data.frame (
  transaction_date = future_dates,
  date_numeric = as.numeric(future_dates),
  day_of_week = wday(future_dates, label = TRUE),
  month = month(future_dates, label = TRUE)
)

future_data$predicted_sales <- predict(svm_model, newdata = future_data)

combined_data <- bind_rows(
  svm_data %>% select(transaction_date, total_sales) %>% mutate(type="Actual"),
           future_data %>% rename(total_sales = predicted_sales) %>% mutate(type = "Predicted")
           )
```

## SVM Model Summary
```{r, echo=FALSE}
summary(svm_model)
```


## Actual and Predicted Coffee Sales (SVM)
```{r, fig.width=7, fig.height=4, echo=FALSE}
ggplot(combined_data, aes(x = transaction_date, y = total_sales, color = type)) +
  geom_line() +
  labs(title = "Actual and Predicted Coffee Sales (SVM)", x = "Date", y = "Sales", color = "Actual and Predicted Coffee Sales") +
  theme_minimal() +
  scale_color_manual(
    values = c("Actual" = "#CB997E", "Predicted" = "#A5A58D")
    )
```

## Predicted total revenue (SVM)
Dec 31, 2023: $1,247,064     
Jan 1, 2023: $1,732.80
```{r, echo=FALSE}
svm_data <- df %>%
  mutate(
    transaction_date = as.Date(transaction_date),
    date_numeric = as.numeric(transaction_date),
    day_of_week = wday(transaction_date, label = TRUE),
    month = month(transaction_date, label = TRUE),
  ) %>%
  group_by(transaction_date) %>%
  summarise(
    total_sales = sum(unit_price, na.rm = TRUE),
    date_numeric = first(date_numeric),
    day_of_week = first(day_of_week),
    month = first(month)
  ) %>%
  ungroup()

svm_model <- svm(total_sales ~ date_numeric + day_of_week + month, data = svm_data, type = "eps-regression")

svm_data$predicted_sales <- predict(svm_model, newdata = svm_data)

last_date <- max(svm_data$transaction_date)
future_dates <- seq.Date(last_date + 1, by = "day", length.out = 184)

future_data <- data.frame (
  transaction_date = future_dates,
  date_numeric = as.numeric(future_dates),
  day_of_week = wday(future_dates, label = TRUE),
  month = month(future_dates, label = TRUE)
)

future_data$predicted_sales <- predict(svm_model, newdata = future_data)

combined_data <- bind_rows(
  svm_data %>% select(transaction_date, total_sales) %>% mutate(type="Actual"),
           future_data %>% rename(total_sales = predicted_sales) %>% mutate(type = "Predicted")
           )

total_revenue <- combined_data %>%
  summarise(expected_total_revenue = sum(total_sales, na.rm = TRUE))
#total_revenue
```




